Supplementary Data: 14C Modeling for Kiwualn Chronology

### 2020-08-21 analysis/paper/KWL-chronology-modeling.Rmd
minor edits when talking about 14C ages

We calibrated 23 dates from different contexts, including normal deposit, burials, and middens, at Kiwulan using oxcAAR package, which executes OxCal v4.4.1 with radiocarbon calibration curve IntCal20 [@Reimer2020].

@@

We then constructed a Bayesian radiocarbon calibration model to obtain the range of pre-European and post-European we examined. We assume the phases are in an order that represents before and after the European arrival in 1624 [@Andrade2007]. The end of post-European is identified by the start of the Chinese migration to Yilan in 1780s [@Chen2007]. The results suggest that pre-European started between 1331–1635 cal AD at a 95.4% probability range and ended between 1458–1670 cal AD at a 99.7% probability range. The end date of pre-European also represents the start date of post-European, while the end date of post-European is 1517-1753 cal AD at a 95.4% probability range. The overall model is consistent with the historical data.

@@

```{r plot}
three_phases_con_results <- executeOxcalScript(three_phases_con)
three_phases_con_text <- readOxcalOutput(three_phases_con_results)
three_phases_result_con_data <- parseOxcalOutput(three_phases_con_text, only.R_Date = F)
str(three_phases_result_con_data)
print(three_phases_result_con_data)
plot(three_phases_result_con_data)
```

>>>>

We calibrated 23 dates from different contexts, including normal deposit, burials, and middens, at Kiwulan using oxcAAR package, which executes OxCal v4.4.1 with radiocarbon calibration curve IntCal20 [@Reimer2020]. <!-- if this is not the full set of ages in Chen, then please explain how you choose them -->

@@

We then constructed a Bayesian radiocarbon calibration model to obtain the range of pre-European and post-European we examined. We assume the phases are in an order that represents before and after the European arrival in 1624 [@Andrade2007]. The end of post-European is identified by the start of the Chinese migration to Yilan in 1780s [@Chen2007].  <!-- please explain how you choose the subset of ages to go into this model -->

The results suggest that the re-European period started between 1331–1635 cal AD at a 95.4% probability range and ended between 1458–1670 cal AD at a 99.7% probability range. The end date of pre-European also represents the start date of post-European, while the end date of post-European is 1517-1753 cal AD at a 95.4% probability range. The overall model is consistent with the historical data.

@@

```{r plot}
three_phases_con_results <- executeOxcalScript(three_phases_con)
three_phases_con_text <- readOxcalOutput(three_phases_con_results)
three_phases_result_con_data <- parseOxcalOutput(three_phases_con_text, only.R_Date = F)
# str(three_phases_result_con_data)
# print(three_phases_result_con_data)
plot(three_phases_result_con_data)
```
